```{r setoptions, echo=FALSE}
opts_chunk$set(warning=FALSE)
```


# Reproducible Research: Peer Assessment 1


```{r}
library(lattice)
```

## Loading and preprocessing the data
```{r}
activity <- read.csv(unz("activity.zip", "activity.csv"))
```

No data preproceesing was required.

## Make a histogram of the total number of steps taken each day

```{r}
agSumStepsDay <- aggregate(steps ~ date, data = activity, sum)
hist(agSumStepsDay$steps, col="cornflowerblue", main="The Total Number of Steps Taken Each Day", xlab="Steps")
```


## What is mean total number of steps taken per day?

```{r}
mean(agSumStepsDay$steps)
```

## What is median total number of steps taken per day?

```{r}
median(agSumStepsDay$steps)
```


## What is the average daily activity pattern?

### Make a time series plot of the 5-minute intervaland the average number of steps taken, averaged across all days

```{r}
agMeanStepsInterval <- aggregate(steps ~ interval, data = activity, mean)
xyplot(steps ~ interval, data=agMeanStepsInterval, type="l",main="average daily activity pattern",xlab="5-minute interval", ylab="number of steps averaged across all days")
```

### Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
agMeanStepsInterval$interval[which(agMeanStepsInterval$steps==max(agMeanStepsInterval$steps))]
```

## Imputing missing values

### Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r}
sum(!complete.cases(activity))
```

### Devise a strategy for filling in all of the missing values in the dataset.

The strategy I chose was to use the mean for the 5-minute interval which the missing value belonged to.

### Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
NoNA<-activity
NARows=which(is.na(NoNA$steps))

for (i in NARows ) {
     NoNA[i,1]<-agMeanStepsInterval[which(agMeanStepsInterval$interval==NoNA[i,3]),2]
 }
```

### Make a histogram of the total number of steps taken each day

```{r}
agSumStepsDayNoNA <- aggregate(steps ~ date, data = NoNA, sum)
hist(agSumStepsDayNoNA$steps, col="cornflowerblue", main="The Total Number of Steps Taken Each Day", xlab="Steps")
```

###  Calculate and report the mean and median total number of steps taken per day

```{r}
mean(agSumStepsDayNoNA$steps)
median(agSumStepsDayNoNA$steps)
```

As we can see, in this example the impact of imputing missing data on the estimates of the total daily number of steps (mean and median) in quite neglectable.


## Are there differences in activity patterns between weekdays and weekends?
```{r}
Sys.setlocale("LC_TIME","English United States")
days<-weekdays(as.Date(NoNA$date))
NoNA$DayType<-as.factor(as.character(factor(days, levels=c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday"),labels=c("weekday","weekday","weekday","weekday","weekday","weekend","weekend"))))
```

Note: the purpose of the use of as.character and as.factor is to make two distinct levels (weekday/weekend) out of seven level (each one for each day of the week). 

```{r}
ag <- aggregate(steps ~ interval+DayType ,data = NoNA, mean)
xyplot(steps ~ interval | DayType, data=ag, type="l",main="average daily activity pattern",xlab="5-minute interval", ylab="number of steps averaged across all days",layout=c(1,3))

```